---
title: "Chicago Updates after Reviews"
author: "Kaylah Thomas (revised from Kate Schertz)"
date: "2/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "~/Box #Sync/Documents/MyOngoing/Crime&Greenspace/AnalysisForRevision/Chicago/")
```



```{r}
library(tidycensus)
library(tidyverse)
library(tigris)
library(rgdal)
library(purrr)
library(sf)
census_api_key("4d745721877aa8f3beca6dd8e074fce5bc74a6ee")

```


Variables needded from ACS 2019 (Kaylah's Version)
*Returns both Estimates and margin of error, formatting to come for percent estimates. Will indicate the total estimate for groups with parameter in get_acs()

Total population: B02001_001
One race-White: B02001_002
One race-Black: B02001_003
One race-American Indian or Alaska Native: B02001_004
One race-Asian:B02001_005
One race-Native Hawaiian: B02001_006
One race-Other: B02001_007
Two or more races:B02001_008
Two races including Some other race: B02001_009
Two races excluding Some other race, and three or more races:B02001_010
Hispanic or Latino: B03001_003
Not Hispanic or Latino: B03001_002
Median Household income: B19013_001
Gini Index: 19083_001
Less than High School Education: B15003_002 to _016
Bachelors or higher: B15003_021 to _025
Total population for age measures: B01001_001
Age range: (have to add male and female populations together)
  Under 5 years: B01001_003 & B01001_027
  5 to 9 years: B01001_004 & B01001_028
  10 to 14 years: B01001_005 & B01001_029
  15 to 17 years: B01001_006 & B01001_030
  18 and 19 years: B01001_007 & B01001_031
  20 years: B01001_008 & B01001_032
  21 years: B01001_009 & B01001_033
  22 to 24 years: B01001_010 & B01001_034
  25 to 29 years: B01001_011 & B01001_035
  30 to 34 years: B01001_012 & B01001_036
  35 to 39 years: B01001_013 & B01001_037
  40 to 44 years: B01001_014 & B01001_038
  45 to 49 years: B01001_015 & B01001_039
  50 to 54 years: B01001_016 & B01001_040
  55 to 59 years: B01001_017 & B01001_041
  60 and 61 years: B01001_018 & B01001_042
  62 to 64 years: B01001_019 & B01001_043
  65 and 66 years: B01001_020 & B01001_044
  67 to 69 years: B01001_021 & B01001_045
  70 to 74 years: B01001_022 & B01001_046
  75 to 79 years: B01001_023 & B01001_047
  80 to 84 years: B01001_024 & B01001_048
  85 years and over: B01001_025 & B01001_049
  * Aggregate all of these age variables to make 7 age ranges: under 18, 18-24, 25-34, 35-44, 45-54, 55-64, 65-74, over 75
Gender: 
  Total population: B01001_001
  Female: B01001_026
  Male: B01001_002
  
  
Adjust age ranges to this? Also included in ACS measures find via B18106_025 (self care difficulty, possible factor to include. Tangent: should inclue geographical mobility in the past year measure?)
18 to 34 years
35 to 64 years
5 to 17 year
65 to 74 years
75 years and over



```{r}
# grabbing FIPS state and county codes for get_acs state and county inputs

cbsa <- read.csv("/Users/kaylah/OneDrive/School Documents/depression_data/list1_Sep_2018.csv")

state_cbsa <- unique(as.character(cbsa[cbsa$X.2 == 'Chicago-Naperville-Elgin, IL-IN-WI', 9]))

county_cbsa <- unique(as.character(cbsa[cbsa$X.2 == 'Chicago-Naperville-Elgin, IL-IN-WI', 8]))

# formattting for multi-city calls to get_acs()
my_counties <- fips_codes %>%
  filter(county %in% county_cbsa) %>%
  filter(state_name %in% state_cbsa)

```

```{r}
# Variables of Interest

race_vars <- c(White = "B03002_003",
               Black = "B03002_004",
               Native = "B03002_005",
               Asian = "B03002_006",
               HIPI = "B03002_007",
               Hispanic = "B03002_012")
race_summ_var <- 'B03002_001'

income_vars <- c(Gini = 'B19083_001', 
                 Medinc = 'B19013_001')

pub_asst_vars <- c(Pubasst = 'B19057_002', 
                   NoPubasst = 'B19057_003')
pub_asst_sum_var <- 'B19057_001'

edu_vars <- c(LessHS1 = 'B15003_002',
              LessHS2 = 'B15003_003',
              LessHS3 = 'B15003_004',
              LessHS4 = 'B15003_005',
              LessHS5 = 'B15003_006',
              LessHS6 = 'B15003_007',
              LessHS7 = 'B15003_008',
              LessHS8 = 'B15003_009',
              LessHS9 = 'B15003_010',
              LessHS10 = 'B15003_011',
              LessHS11 = 'B15003_012',
              LessHS12 = 'B15003_013',
              LessHS13 = 'B15003_014',
              LessHS14 = 'B15003_015',
              LessHS15 = 'B15003_016',
              College1 = 'B15003_021',
              College2 = 'B15003_022',
              College3 = 'B15003_023',
              College4 = 'B15003_024',
              College5 = 'B15003_025'
              )
edu_sum_var <- 'B15003_001'

age_vars <- c(mUnder5 = 'B01001_003', 
            fUnder5 = 'B01001_027',
            m5t9 = 'B01001_004',
            f5t9 = 'B01001_028',
            m10t14 = 'B01001_005',
            f10t14 = 'B01001_029',
            m15t17 = 'B01001_006',
            f15t17 = 'B01001_030',
            m18t19 = 'B01001_007',
            f18t19 = 'B01001_031',
            m20 = 'B01001_008',
            f20 = 'B01001_032',
            m21 = 'B01001_009',
            f21 = 'B01001_033',
            m22t24 = 'B01001_010',
            f22t24 = 'B01001_034',
            m25t29 = 'B01001_011',
            f25t29 = 'B01001_035',
            m30t34 = 'B01001_012',
            f30t34 = 'B01001_036',
            m35t39 = 'B01001_013',
            f35t39 = 'B01001_037',
            m40t44 = 'B01001_014',
            f40t44 = 'B01001_038',
            m45t49 = 'B01001_015',
            f45t49 = 'B01001_039',
            m50t54 = 'B01001_016',
            f50t54 = 'B01001_040',
            m55t59 = 'B01001_017',
            f55t59 = 'B01001_041',
            m60t61 = 'B01001_018',
            f60t61 = 'B01001_042',
            m62t64 = 'B01001_019',
            f62t64 = 'B01001_043',
            m65t66 = 'B01001_020',
            f65t66 = 'B01001_044',
            m67t69 = 'B01001_021',
            f67t69 = 'B01001_045',
            m70t74 = 'B01001_022',
            f70t74 = 'B01001_046',
            m75t79 = 'B01001_023',
            f75t79 = 'B01001_047',
            m80t84 = 'B01001_024',
            f80t84 = 'B01001_048',
            m85over = 'B01001_025',
            f85over = 'B01001_049')
age_pop_tot <- 'B01001_001'

gender_vars <- c(female = 'B01001_026',
                 male = 'B01001_002')
gender_tot <- 'B01001_001'
```


```{r}
#Kaylah
# have to use acs 2019 because relevant variables for 2020 census have not been released yet

v19 <- load_variables(2019, "acs5", cache = TRUE)
View(v19)

# Race Variables
chi_cbsa_race <- map2_dfr(
  my_counties$state_code, my_counties$county_code,
  ~ get_acs(geography = "tract", state = .x, county = .y, 
            variables = race_vars,
            summary_var = race_summ_var,
            year = 2019, 
            output = "wide")
  )

chi_cbsa_race_sum <- chi_cbsa_race %>%
  rowwise(GEOID) %>%
  mutate(PEWhite = WhiteE/summary_est,
         PEBlack = BlackE/summary_est,
         PENative = NativeE/summary_est,
         PEAsian = AsianE/summary_est,
         PEHIPI = HIPIE/summary_est, 
         PEHispanic = HispanicE/summary_est)
chi_cbsa_race_sum <- chi_cbsa_race_sum %>% rename(RaceSum = summary_est, RaceMOE = summary_moe)


# Income and Gini index Variables
chi_cbsa_inc <- map2_dfr(
  my_counties$state_code, my_counties$county_code,
  ~ get_acs(geography = "tract", state = .x, county = .y, 
            variables = income_vars,
            year = 2019, 
            output = "wide")
  ) # Note: income variables have 17 (max) Nan values


# Public Assitance Variables
chi_cbsa_pubasst <- map2_dfr(
  my_counties$state_code, my_counties$county_code,
  ~ get_acs(geography = "tract", state = .x, county = .y, 
            variables = pub_asst_vars,
            summary_var = pub_asst_sum_var, 
            year = 2019, 
            output = "wide")
  )

chi_cbsa_pubasst_sum <- chi_cbsa_pubasst %>%
  rowwise(GEOID) %>%
  mutate(PEpubasst = PubasstE/summary_est,
         PEnopubasst = NoPubasstE/summary_est)
chi_cbsa_pubasst_sum <- chi_cbsa_pubasst_sum %>% rename(PubAsstSum = summary_est, PubAsstMOE = summary_moe)


# Education variables
chi_cbsa_edu <- map2_dfr(
  my_counties$state_code, my_counties$county_code,
  ~ get_acs(geography = "tract", state = .x, county = .y, 
            variables = edu_vars,
            summary_var = edu_sum_var, 
            year = 2019, 
            output = "wide")
  )

# Only have less than HS and bachelor or higher, can add complete highschool and some college vars later 
chi_cbsa_edu_sum <- chi_cbsa_edu %>%
  rowwise(GEOID) %>%
  mutate(PENoCollege = sum(c(LessHS1E,LessHS2E, LessHS3E, 
                           LessHS4E, LessHS5E, LessHS6E, 
                           LessHS7E, LessHS8E, LessHS9E, 
                           LessHS10E, LessHS11E, LessHS12E,
                           LessHS13E, LessHS14E, LessHS15E))/summary_est,
         PECollege = (sum(College1E, College2E, College3E, College4E, College5E))/summary_est)
chi_cbsa_edu_sum <- chi_cbsa_edu_sum %>% rename(EdutSum = summary_est, EduMOE = summary_moe)

# Age range variables
chi_cbsa_age <- map2_dfr(
  my_counties$state_code, my_counties$county_code,
  ~ get_acs(geography = "tract", state = .x, county = .y, 
            variables = age_vars,
            summary_var = age_pop_tot, 
            year = 2019, 
            output = "wide")
  )
# Merging Female and Male age ranges together
#chi_cbsa_age_sum <- chi_cbsa_age %>%
#  rowwise(GEOID) %>%
#  mutate(PEUnder5 = sum(c(mUnder5E, 
#                         fUnder5E))/summary_est,
#         PE5t9 = sum(c(m5t9E,
#                      f5t9E))/summary_est,
#         PE10t14 = sum(c(m10t14E,
#                        f10t14E))/summary_est,
#         PE15t17 = sum(c(m15t17E,
#                        f15t17E))/summary_est,
#         PE18t19 = sum(c(m18t19E,
#                        f18t19E))/summary_est,
#         PE20 = sum(c(m20E,
#                     f20E))/summary_est,
#         PE21 = sum(c(m21E,
#                     f21E))/summary_est,
#         PE22t24 = sum(c(m22t24E,
#                        f22t24E))/summary_est,
#         PE25t29 = sum(c(m25t29E,
#                        f25t29E))/summary_est,
#         PE30t34 = sum(c(m30t34E,
#                        f30t34E))/summary_est,
#         PE35t39 = sum(c(m35t39E,
#                        f35t39E))/summary_est,
#         PE40t44 = sum(c(m40t44E,
#                        f40t44E))/summary_est,
#         PE50t54 = sum(c(m50t54E,
#                        f50t54E))/summary_est,
#         PE55t59 = sum(c(m55t59E,
#                        f55t59E))/summary_est,
#         PE60t61 = sum(c(m60t61E,
#                        f60t61E))/summary_est,
#         PE62t64 = sum(c(m62t64E,
#                        f62t64E))/summary_est,
#         PE65t66 = sum(c(m65t66E,
#                        f65t66E))/summary_est,
#         PE67t69 = sum(c(m67t69E,
#                        f67t69E))/summary_est,
#         PE70t74 = sum(c(m70t74E,
#                        f70t74E))/summary_est,
#         PE75t79 = sum(c(m75t79E,
#                        f75t79E))/summary_est,
#         PE80t84 = sum(c(m80t84E,
#                        f80t84E))/summary_est,
#         PE85over = sum(c(m85overE,
#                         f85overE))/summary_est)

chi_cbsa_age_sum <- chi_cbsa_age %>%
  rowwise(GEOID) %>%
  mutate(PEUnder18 = sum(c(mUnder5E,
                           fUnder5E,
                           m5t9E,
                           f5t9E,
                           m10t14E,
                           f10t14E,
                           m15t17E,
                           f15t17E))/summary_est,
         PE18t24 = sum(c(m18t19E,
                         f18t19E,
                         m20E,
                         f20E,
                         m21E,
                         f21E,
                         m22t24E,
                         f22t24E))/summary_est,
         PE25t34 = sum(c(m25t29E,
                        f25t29E,
                        m30t34E,
                        f30t34E))/summary_est,
         PE35t34 = sum(c(m35t39E,
                          f35t39E,
                          m40t44E,
                          f40t44E))/summary_est,
         PE45t54 = sum(c(m45t49E,
                         f45t49E,
                         m50t54E,
                         f50t54E))/summary_est,
         PE55t64 = sum(c(m55t59E,
                          f55t59E,
                          m60t61E,
                          f60t61E,
                          m62t64E,
                          f62t64E))/summary_est,
         PE65t74 = sum(c(m65t66E,
                         f65t66E,
                         m67t69E,
                         f67t69E,
                         m70t74E,
                         f70t74E))/summary_est,
         PE75over = sum(c(m75t79E,
                          f75t79E,
                          m80t84E,
                          f80t84E,
                          m85overE,
                          f85overE))/summary_est)
chi_cbsa_age_sum <- chi_cbsa_age_sum %>% rename(AgeSum = summary_est, AgeMOE = summary_moe)

# Gender Variables
chi_cbsa_gender <- map2_dfr(
  my_counties$state_code, my_counties$county_code,
  ~ get_acs(geography = "tract", state = .x, county = .y, 
            variables = gender_vars,
            summary_var = gender_tot,
            year = 2019, 
            output = "wide")
  )

chi_cbsa_gender_sum <- chi_cbsa_gender %>%
  rowwise(GEOID) %>%
  mutate(PEFemale = femaleE/summary_est,
         PEMale = maleE/summary_est)
chi_cbsa_gender_sum <- chi_cbsa_gender_sum %>% rename(GenderSum = summary_est, GenderMOE = summary_moe)



# Merging and Cleaning Dataset
Chi_acs2019 <- merge(chi_cbsa_race_sum, chi_cbsa_edu_sum, by="GEOID")
Chi_acs2019 <- merge(Chi_acs2019, chi_cbsa_pubasst_sum, by='GEOID')
Chi_acs2019 <- merge(Chi_acs2019, chi_cbsa_inc, by='GEOID')
Chi_acs2019 <- merge(Chi_acs2019, chi_cbsa_age_sum, by='GEOID')
Chi_acs2019 <- merge(Chi_acs2019, chi_cbsa_gender_sum, by='GEOID')



# Chi_acs2019 <- Chi_acs2019 %>% select(-23,-68, -77, -82, -200)
Chi_acs2019 <- Chi_acs2019 %>% select(-23,-68, -77, -82, -185)
Chi_acs2019 <- Chi_acs2019 %>% rename(Tract = 2)

write_csv(Chi_acs2019, 'Chicago_subset_acs2019.csv')

# Chi_acs2019_clean <- Chi_acs2019 %>% 
#  select(GEOID, Tract, PEWhite, PEBlack, PENative, PEAsian, PEHIPI, PEHispanic, GiniE, MedincE, PEpubasst,
#         PEnopubasst,PENoCollege, PECollege, PEUnder5, PE5t9, PE10t14, PE15t17, PE18t19, PE20, PE21, PE22t24, #PE25t29, PE30t34, PE35t39, PE40t44, PE45t49, PE50t54, PE55t59, PE60t61, PE62t64, PE65t66, PE67t69, PE70t74, #PE75t79, PE80t84, PE85over, PEFemale, PEMale)

Chi_acs2019_clean <- Chi_acs2019 %>% 
  select(GEOID, Tract, PEWhite, PEBlack, PENative, PEAsian, PEHIPI, PEHispanic, GiniE, MedincE, PEpubasst,
         PEnopubasst,PENoCollege, PECollege, PEUnder18, PE18t24, PE25t34, PE35t34, PE45t54, PE55t64, PE65t74,
         PE75over, PEFemale, PEMale)

write_csv(Chi_acs2019_clean,"Chicago_subset_acs2019_clean.csv")
```

Building custom shapefile
```{r}
multi_state_ct_geo_list <- map2(
  my_counties$state_code, my_counties$county_code,
    ~ get_acs(
        geography = "tract",
        variables = c(pop_tot = race_summ_var),
        state = .x,
        county = .y,
        year = 2019,
        geometry = TRUE,
        output = "wide"  # get data in wide format for easier mapping
        )
    )

multi_state_ct_geo <- reduce(multi_state_ct_geo_list, rbind)

st_write(multi_state_ct_geo, "custom_chicago_cbsa_tracts.shp", append=FALSE)
```


Modifying chunk to use zipcode for prelim analysis, not tract 
```{r}
#Kaylah
# have to use acs 2019 because relevant variables for 2020 census have not been released yet

v19 <- load_variables(2019, "acs5", cache = TRUE)
View(v19)

State_lst <- c('IN', 'IL', 'WI')

# Race Variables
chi_cbsa_race <- get_acs(geography = "zcta", state = State_lst, 
            variables = race_vars,
            summary_var = race_summ_var,
            year = 2019, 
            output = "wide")

chi_cbsa_race_sum <- chi_cbsa_race %>%
  rowwise(GEOID) %>%
  mutate(PEWhite = WhiteE/summary_est,
         PEBlack = BlackE/summary_est,
         PENative = NativeE/summary_est,
         PEAsian = AsianE/summary_est,
         PEHIPI = HIPIE/summary_est, 
         PEHispanic = HispanicE/summary_est)
chi_cbsa_race_sum <- chi_cbsa_race_sum %>% rename(RaceSum = summary_est, RaceMOE = summary_moe)


# Income and Gini index Variables
chi_cbsa_inc <- get_acs(geography = "zcta", state = State_lst,
            variables = income_vars,
            year = 2019, 
            output = "wide") # Note: income variables have 17 (max) Nan values


# Public Assitance Variables
chi_cbsa_pubasst <- get_acs(geography = "zcta", state = State_lst, 
            variables = pub_asst_vars,
            summary_var = pub_asst_sum_var, 
            year = 2019, 
            output = "wide")

chi_cbsa_pubasst_sum <- chi_cbsa_pubasst %>%
  rowwise(GEOID) %>%
  mutate(PEpubasst = PubasstE/summary_est,
         PEnopubasst = NoPubasstE/summary_est)
chi_cbsa_pubasst_sum <- chi_cbsa_pubasst_sum %>% rename(PubAsstSum = summary_est, PubAsstMOE = summary_moe)


# Education variables
chi_cbsa_edu <- get_acs(geography = "zcta", state = State_lst,
            variables = edu_vars,
            summary_var = edu_sum_var, 
            year = 2019, 
            output = "wide")

# Only have less than HS and bachelor or higher, can add complete highschool and some college vars later 
chi_cbsa_edu_sum <- chi_cbsa_edu %>%
  rowwise(GEOID) %>%
  mutate(PENoCollege = sum(c(LessHS1E,LessHS2E, LessHS3E, 
                           LessHS4E, LessHS5E, LessHS6E, 
                           LessHS7E, LessHS8E, LessHS9E, 
                           LessHS10E, LessHS11E, LessHS12E,
                           LessHS13E, LessHS14E, LessHS15E))/summary_est,
         PECollege = (sum(College1E, College2E, College3E, College4E, College5E))/summary_est)
chi_cbsa_edu_sum <- chi_cbsa_edu_sum %>% rename(EdutSum = summary_est, EduMOE = summary_moe)

# Age range variables
chi_cbsa_age <- get_acs(geography = "zcta", state = State_lst, 
            variables = age_vars,
            summary_var = age_pop_tot, 
            year = 2019, 
            output = "wide")

# Merging Female and Male age ranges together
chi_cbsa_age_sum <- chi_cbsa_age %>%
  rowwise(GEOID) %>%
    mutate(PEUnder18 = sum(c(mUnder5E,
                           fUnder5E,
                           m5t9E,
                           f5t9E,
                           m10t14E,
                           f10t14E,
                           m15t17E,
                           f15t17E))/summary_est,
         PE18t24 = sum(c(m18t19E,
                         f18t19E,
                         m20E,
                         f20E,
                         m21E,
                         f21E,
                         m22t24E,
                         f22t24E))/summary_est,
         PE25t34 = sum(c(m25t29E,
                        f25t29E,
                        m30t34E,
                        f30t34E))/summary_est,
         PE35t34 = sum(c(m35t39E,
                          f35t39E,
                          m40t44E,
                          f40t44E))/summary_est,
         PE45t54 = sum(c(m45t49E,
                         f45t49E,
                         m50t54E,
                         f50t54E))/summary_est,
         PE55t64 = sum(c(m55t59E,
                          f55t59E,
                          m60t61E,
                          f60t61E,
                          m62t64E,
                          f62t64E))/summary_est,
         PE65t74 = sum(c(m65t66E,
                         f65t66E,
                         m67t69E,
                         f67t69E,
                         m70t74E,
                         f70t74E))/summary_est,
         PE75over = sum(c(m75t79E,
                          f75t79E,
                          m80t84E,
                          f80t84E,
                          m85overE,
                          f85overE))/summary_est)
chi_cbsa_age_sum <- chi_cbsa_age_sum %>% rename(AgeSum = summary_est, AgeMOE = summary_moe)

# Gender Variables
chi_cbsa_gender <- get_acs(geography = "zcta", state = State_lst, 
            variables = gender_vars,
            summary_var = gender_tot,
            year = 2019, 
            output = "wide")

chi_cbsa_gender_sum <- chi_cbsa_gender %>%
  rowwise(GEOID) %>%
  mutate(PEFemale = femaleE/summary_est,
         PEMale = maleE/summary_est)
chi_cbsa_gender_sum <- chi_cbsa_gender_sum %>% rename(GenderSum = summary_est, GenderMOE = summary_moe)



# Merging and Cleaning Dataset
Chi_acs2019 <- merge(chi_cbsa_race_sum, chi_cbsa_edu_sum, by="GEOID")
Chi_acs2019 <- merge(Chi_acs2019, chi_cbsa_pubasst_sum, by='GEOID')
Chi_acs2019 <- merge(Chi_acs2019, chi_cbsa_inc, by='GEOID')
Chi_acs2019 <- merge(Chi_acs2019, chi_cbsa_age_sum, by='GEOID')
Chi_acs2019 <- merge(Chi_acs2019, chi_cbsa_gender_sum, by='GEOID')


# Chi_acs2019 <- Chi_acs2019 %>% select(-23,-68, -77, -82, -200)
Chi_acs2019 <- Chi_acs2019 %>% select(-23,-68, -77, -82, -185)
Chi_acs2019 <- Chi_acs2019 %>% rename(Tract = 2)

write_csv(Chi_acs2019, 'Chicago_zcta_subset_acs2019.csv')

Chi_acs2019_clean <- Chi_acs2019 %>% 
  select(GEOID, Tract, PEWhite, PEBlack, PENative, PEAsian, PEHIPI, PEHispanic, GiniE, MedincE, PEpubasst,
         PEnopubasst,PENoCollege, PECollege, PEUnder18, PE18t24, PE25t34, PE35t34, PE45t54, PE55t64, PE65t74,
         PE75over, PEFemale, PEMale)

write_csv(Chi_acs2019_clean,"Chicago_zcta_subset_acs2019_clean.csv")
```


Merging ACS variables and crime, and reducing from Cook to Chicago census tracts
```{r}
Chicago_tracts <- read_csv("../../Chicago_DataSetupFiles/CensusTracts_Chicago.csv")
Chicago_tracts <- Chicago_tracts %>% select(geoid10) %>% mutate(geoid10=as.factor(geoid10)) %>% rename(GEOID=geoid10)
Chicago_2017variables <- merge(Chicago_tracts, cleanvars, all.x = TRUE, by="GEOID")

chi_crime2017 <- read.csv("ChicagoCrimeByCensusTract_2017.csv")
Chicago_2017variables <- merge(Chicago_2017variables, chi_crime2017, all.x = TRUE)

write_csv(Chicago_2017variables, "Chicago2017_ACS&Crime.csv")
```

Adding working population
From https://lehd.ces.census.gov/data/
LODES Survey - WAC characteristics
C000 is total number of jobs, starts at Census Block level
Need to group by census tract (first 11 digit)
```{r}
il_wac <- read_csv("il_wac_S000_JT00_2017.csv")
cook_wac <- il_wac %>% filter(grepl("^17031", w_geocode)) %>% dplyr::select(w_geocode, C000)
cook_wac$w_geocode <- as.character(cook_wac$w_geocode)
cook_wac <- cook_wac %>% mutate(tract = substr(w_geocode, 1, 11))
cook_wac$tract <- as.factor(cook_wac$tract)
cook_wac_ct <- cook_wac %>% group_by(tract) %>% dplyr::summarize(WorkingPop = sum(C000))

#Adding to all other variables
Chicago_2017variables <- merge(Chicago_2017variables, cook_wac_ct, all.x=TRUE, by.x="GEOID", by.y="tract")

#If a census tract is missing it means they had no jobs there in this year, so changing to 0
Chicago_2017variables <- Chicago_2017variables %>% 
  replace_na(list(WorkingPop=0))

write_csv(Chicago_2017variables, "Chicago2017_ACS&Crime.csv")
```

Getting Land Cover from original data
```{r}
originaldata <- read_csv("../../TractsChicago_withAllData.csv")
orig_neededvars <- originaldata %>% dplyr::select(GEOID, CommunityArea, PercTreeZScore, PercGrassZScore, knnZScore, knn, PercTrees, PercGrass)

Chicago_2017variables <- merge(orig_neededvars, Chicago_2017variables, all.x = TRUE)

summary(Chicago_2017variables)
colnames(Chicago_2017variables)

write_csv(Chicago_2017variables,"Chicago2017_ACS&Crime&Land.csv")
```

Adding Jamie's new Park Visits data
```{r}
newpark_mus <- read.csv("chicago_psych_100m.csv")
newpark_mus <- newpark_mus %>% rename(GEOID=geoid) %>% mutate(GEOID=as.factor(GEOID))
newpark_mus <- newpark_mus %>% rename(NewPark = npark, NewMuseum = nmuse)

newpark_dist <- read.csv("chicago_park_dist_100m.csv")
newpark_dist <- newpark_dist %>% rename(GEOID=geoid) %>% mutate(GEOID=as.factor(GEOID))

Chicago_2017variables <- merge(Chicago_2017variables, newpark_mus, all.x = TRUE, by= "GEOID")
Chicago_2017variables <- merge(Chicago_2017variables, newpark_dist, all.x = TRUE, by= "GEOID")

write_csv(Chicago_2017variables, "Chicago2017_ACS&Crime&Land&Cell.csv")
```

Z-scoring and Log Transformations
```{r}
Chicago_2017variables$GEOID <- as.factor(Chicago_2017variables$GEOID)
Chicago_2017variables$CommunityArea <- as.factor(Chicago_2017variables$CommunityArea)

Chicago_2017variables <- Chicago_2017variables %>% 
  mutate(PercBlackAloneZScore = (PercBlackAlone - mean(PercBlackAlone))/sd(PercBlackAlone),
         PercBlackOtherZScore = (PercBlackAloneWOther - mean(PercBlackAloneWOther))/sd(PercBlackAloneWOther),
         PercHispanicZScore = (PercHispanic - mean(PercHispanic))/sd(PercHispanic),
         PercCollegeZScore = (PercCollege - mean(PercCollege))/sd(PercCollege),
         PercLessThanHSZScore = (PercLessThanHS - mean(PercLessThanHS))/sd(PercLessThanHS),
         PercPovertyZScore = (PercPoverty - mean(PercPoverty))/sd(PercPoverty),
         PercUnemployedZScore = (PercUnemployed - mean(PercUnemployed))/sd(PercUnemployed),
         PercResStabZScore = (PercResStability - mean(PercResStability))/sd(PercResStability),
         PercForeignBornZScore = (PercForeignBorn - mean(PercForeignBorn))/sd(PercForeignBorn),
         PercCrowdedZScore = (PercCrowded - mean(PercCrowded))/sd(PercCrowded),
         LogMedHouseInc = log(MedHouseInc),
         LogIncomeZScore = (LogMedHouseInc - mean(LogMedHouseInc))/sd(LogMedHouseInc),
         LogNonViolent = log(NonViolent),
         LogViolent = log(Violent),
         TotalCrime = NonViolent + Violent,
         LogWorkingPop = log1p(WorkingPop),
         LogJobsZScore = (LogWorkingPop - mean(LogWorkingPop))/sd(LogWorkingPop),
         LogPopulation = log(population),
         LogPopZScore = (LogPopulation - mean(LogPopulation))/sd(LogPopulation),
         NewParkZScore = (NewPark - mean(NewPark))/sd(NewPark),
         NewMuseumZScore = (NewMuseum - mean(NewMuseum))/sd(NewMuseum),
         parkdistZScore = (tp_dist - mean(tp_dist))/sd(tp_dist))

#Checking for any missing data
sum(is.na.data.frame(Chicago_2017variables))

VarsForModel <- Chicago_2017variables %>% 
  dplyr::select(GEOID, CommunityArea, LogNonViolent, LogViolent, PercTreeZScore, PercGrassZScore, knnZScore, parkdistZScore, NewParkZScore, NewMuseumZScore, PercBlackOtherZScore, PercHispanicZScore, PercCollegeZScore, PercLessThanHSZScore, PercPovertyZScore, PercUnemployedZScore, PercResStabZScore, PercForeignBornZScore, PercCrowdedZScore, LogIncomeZScore, LogJobsZScore, LogPopZScore)


write_csv(Chicago_2017variables, "Chicago2017_AllData.csv")

#Read this back in to start analysis. Will need 'all variables' version if interpretation of beta
write_csv(VarsForModel, "Chicago2017_VarsForModel.csv")
```

Land Use data
```{r}
landuse <- readOGR("../Parcel_Chicago/.")
```

